---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ppvsbp

<!-- badges: start -->
<!-- badges: end -->

The aim of `ppvsbp` package is to make available the code used in the paper
" A comparison between precise and bounded probability to quantify epistemic uncertainty
 in a scientific assessment"
 
This is a simplified version of an existing scientific assessment on the weekly intake dose of aluminium via consumption of cocoa and chocolate products done by BfR. 
The aim of the assessment was to estimate the 2017 average long-term aluminium intake (chronic toxicity) by consumption of chocolate and cocoa products in 2017 for infants from age 0.5 years to less than 5 years (which are not breastfed) in Germany for the 95th percentile of the specified population  (in $\mu$g/(kg bw)/week). The results were compared with the tolerable weekly intake (TWI) of 1 mg/kg for aluminium.

This is a common type of scientific assessment where the assessment model contains several variables which are combined to estimate the probability that a child exceeds the tolerable weekly intake of aluminium from consuming chocolate products. 
...

## Installation

You can install the development version of `ppvsbp` from [github](https://github.com/Iraices/PrecisePvsBoundedP) with:

``` r
remotes::install_github('Iraices/PrecisePvsBoundedP')
```

## Example
This example does the aluminium exposure assessment by precise probability derived by Bayesian analysis.

First, let us estimate the frequency of exceeding the safety threshold for a random child

```{r}
library(ppvsbp)
## basic example code
TWI_pp_random_child <- 
  unc_analysis_assessment(niter_ale = 5000, niter_epi = 5000, threshold = 1,
                          percentile_ale = NULL,
                          suff_stat_concentration = data_assessment$log_concentration_ss_data,
                          suff_stat_consumption = data_assessment$log_consumption_ss_data,
                          consumption_change_vals_EKE = data_assessment$change_cons$vals,
                          consumption_change_probs_EKE = data_assessment$change_cons$probs/100,
                          consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                          concentration_mu0 = 3.5, concentration_v0 = 5, concentration_alpha0 = 1,
                          concentration_beta0 = 1, sufficient_statistics_concentration = TRUE,
                          consumption_mu0 = -3, consumption_v0 = 5, consumption_alpha0 = 1,
                          consumption_beta0 = 1, sufficient_statistics_consumption = TRUE,
                          consumption_event_alpha0 = 1, consumption_event_beta0 = 1)

## Expected frequency of exceeding the safety threshold
TWI_pp_random_child$expected_frequency_exceeding

## Highest posterior density interval of the frequency of exceeding the safety threshold
TWI_pp_random_child$hdi_frequency_exceeding

```

Second, let us estimate the frequency of exceeding the safety threshold for a high consumer child. Note, the argument
`percentile_ale` has changed `percentile_ale = 95`.

```{r}
TWI_pp_high_consumer_child <- 
  unc_analysis_assessment(niter_ale = 5000, niter_epi = 5000, threshold = 1,
                          percentile_ale = 95,
                          suff_stat_concentration = data_assessment$log_concentration_ss_data,
                          suff_stat_consumption = data_assessment$log_consumption_ss_data,
                          consumption_change_vals_EKE = data_assessment$change_cons$vals,
                          consumption_change_probs_EKE = data_assessment$change_cons$probs/100,
                          consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                          concentration_mu0 = 3.5, concentration_v0 = 5, concentration_alpha0 = 1,
                          concentration_beta0 = 1, sufficient_statistics_concentration = TRUE,
                          consumption_mu0 = -3, consumption_v0 = 5, consumption_alpha0 = 1,
                          consumption_beta0 = 1, sufficient_statistics_consumption = TRUE,
                          consumption_event_alpha0 = 1, consumption_event_beta0 = 1)

## Expected probability of exceeding the safety threshold
TWI_pp_high_consumer_child$expected_frequency_exceeding

## Highest posterior density interval of the probability of exceeding the safety threshold
TWI_pp_high_consumer_child$hdi_frequency_exceeding
```

We can now visualize the results. Epistemic uncertainty has been characterized by a full probability distribution.

```{r fig1, out.width = '70%'}
 ## A random children
  graph_pp(frequency_exceeding = TWI_pp_random_child$frequency_exceeding)
    
 ## A high consumer child
  graph_pp(frequency_exceeding = TWI_pp_high_consumer_child$frequency_exceeding)
    
```

```{r fig2, out.width = '85%'}
 ## both graphs
 graph_pp_both(frequency_exceeding_random_child = TWI_pp_random_child$frequency_exceeding, 
               frequency_exceeding_high_consumer_child = TWI_pp_high_consumer_child$frequency_exceeding)
```

Now, let us estimate the frequency of exceeding the safety threshold by bounded probability derived by robust Bayesian analysis. 

First, let us consider the case of a random child

```{r}
## A random child
lower_bound_random_child <- 
  bound_frequency_exceeding_bp(obj_func_bp = obj_func_bp, maximize = FALSE, 
                               lower_parameters  = c(1, -5, -20), 
                               upper_parameters  = c(6, 1, -10),
                               niter_ale = 2000, niter_epi = 2000, threshold = 1, percentile_ale = NULL,
                               suff_stat_concentration = data_assessment$log_concentration_ss_data,
                               suff_stat_consumption = data_assessment$log_consumption_ss_data,
                               consumption_change_vals_EKE = c(-15, 7.5), 
                               consumption_change_probs_EKE = c(0.25, 0.75),
                               consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                               concentration_mu0 = 2.75,
                               concentration_v0 = 5, concentration_alpha0 = 1, concentration_beta0 = 1, 
                               sufficient_statistics_concentration = TRUE,
                               consumption_mu0 = -2.5,
                               consumption_v0 = 5, consumption_alpha0 = 1, consumption_beta0 = 1, 
                               sufficient_statistics_consumption = TRUE,
                               consumption_event_alpha0 = 1, consumption_event_beta0 = 1)
  

## A random child
upper_bound_random_child <- 
  bound_frequency_exceeding_bp(obj_func_bp = obj_func_bp, maximize = TRUE, 
                               lower_parameters  = c(1, -5, -20), 
                               upper_parameters  = c(6, 1, -10),
                               niter_ale = 2000, niter_epi = 2000, threshold = 1, 
                               percentile_ale = NULL,
                               suff_stat_concentration = data_assessment$log_concentration_ss_data,
                               suff_stat_consumption = data_assessment$log_consumption_ss_data,
                               consumption_change_vals_EKE = c(-15, 7.5), 
                               consumption_change_probs_EKE = c(0.25, 0.75),
                               consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                               concentration_mu0 = 2.75,
                               concentration_v0 = 5, concentration_alpha0 = 1, concentration_beta0 = 1, 
                               sufficient_statistics_concentration = TRUE,
                               consumption_mu0 = -2.5,
                               consumption_v0 = 5, consumption_alpha0 = 1, consumption_beta0 = 1, 
                               sufficient_statistics_consumption = TRUE,
                               consumption_event_alpha0 = 1, consumption_event_beta0 = 1)
  
```

Now, let us consider the case of a high consumer child.

```{r}
  ## A high consumer child
  lower_bound_high_consumer_child <- 
    bound_frequency_exceeding_bp(obj_func_bp = obj_func_bp, maximize = FALSE, 
                                 lower_parameters  = c(1, -5, -20), 
                                 upper_parameters  = c(6, 1, -10),
                                 niter_ale = 2000, niter_epi = 2000, threshold = 1, 
                                 percentile_ale = 95,
                                 suff_stat_concentration = data_assessment$log_concentration_ss_data,
                                 suff_stat_consumption = data_assessment$log_consumption_ss_data,
                                 consumption_change_vals_EKE = c(-15, 7.5), 
                                 consumption_change_probs_EKE = c(0.25, 0.75),
                                 consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                                 concentration_mu0 = 2.75,
                                 concentration_v0 = 5, concentration_alpha0 = 1, 
                                 concentration_beta0 = 1, 
                                 sufficient_statistics_concentration = TRUE,
                                 consumption_mu0 = -2.5,
                                 consumption_v0 = 5, consumption_alpha0 = 1, consumption_beta0 = 1, 
                                 sufficient_statistics_consumption = TRUE,
                                 consumption_event_alpha0 = 1, consumption_event_beta0 = 1)
  
  
  
## A high consumer child
upper_bound_high_consumer_child <- 
    bound_frequency_exceeding_bp(obj_func_bp = obj_func_bp, maximize = TRUE, 
                                 lower_parameters  = c(1, -5, -20), 
                                 upper_parameters  = c(6, 1, -10),
                                 niter_ale = 2000, niter_epi = 2000, threshold = 1, 
                                 percentile_ale = 95,
                                 suff_stat_concentration = data_assessment$log_concentration_ss_data,
                                 suff_stat_consumption = data_assessment$log_consumption_ss_data,
                                 consumption_change_vals_EKE = c(-15, 7.5), 
                                 consumption_change_probs_EKE = c(0.25, 0.75),
                                 consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                                 concentration_mu0 = 2.75,
                                 concentration_v0 = 5, concentration_alpha0 = 1, 
                                 concentration_beta0 = 1, 
                                 sufficient_statistics_concentration = TRUE,
                                 consumption_mu0 = -2.5,
                                 consumption_v0 = 5, consumption_alpha0 = 1, consumption_beta0 = 1, 
                                 sufficient_statistics_consumption = TRUE,
                                 consumption_event_alpha0 = 1, consumption_event_beta0 = 1)
```

In the case of bounded probability derived by robust Bayesian analysis, epistemic uncertainty has been characterized by a
probabilty box. The results of both cases can be visualized individualy or together.

Individual cases (a random child and a high consumer child)

```{r fig3, out.width = '70%'}
 ## All children
graph_bp(lower_points = lower_bound_random_child$opt_freq$frequency_exceeding, 
         upper_points = upper_bound_random_child$opt_freq$frequency_exceeding)
    
 ## A high consumer child
graph_bp(lower_points = lower_bound_high_consumer_child$opt_freq$frequency_exceeding, 
         upper_points = upper_bound_high_consumer_child$opt_freq$frequency_exceeding)
```

Both cases together (a random child and a high consumer child)

```{r fig4, out.width = '70%'}  
 ## Both graphs
graph_bp_both(lower_points_random_child = lower_bound_random_child$opt_freq$frequency_exceeding, 
              upper_points_random_child = upper_bound_random_child$opt_freq$frequency_exceeding,
              lower_points_high_consumer_child = lower_bound_high_consumer_child$opt_freq$frequency_exceeding, 
              upper_points_high_consumer_child = upper_bound_high_consumer_child$opt_freq$frequency_exceeding)
```

As an example:
Let us assume that a decision maker wants to know if the frequency of exceeding the safety threshold is low.
Let us set 0.05, as an acceptable level of frequency of exceeding.

Thus, the decision maker is interested in $$P(h > 0.05)$$ is low.

In the precise probability case, the probability that the frequency of exceeding the safety threshold of a random child exceeds the acceptable level is

$$P(h >= 0.05) = 1 - P(h <= 0.05) = \text{prob}$$ 

```{r}

index =  which(sort(TWI_pp_random_child$frequency_exceeding)>= 0.05)[1]

val <- function(index, frequency_exceeding){
  if(is.na(index)){
  bound <- 0
  }
  else{
  bound <- 1 - index / length(frequency_exceeding)
  }
  bound
}

prob = val(index = index, frequency_exceeding = TWI_pp_random_child$frequency_exceeding)
prob

```

In the bounded probability case, the probability that the frequency of exceeding the safety threshold of a random child exceeds the acceptable level is

$$P(h >= 0.05) = 1 - P(h <= 0.05)$$ 

$$lower\_bound  <= 1 - P(h <= 0.05)  <= upper\_bound$$ 

```{r}

index_lower <- which(sort(lower_bound_random_child$opt_freq$frequency_exceeding)>= 0.05)[1]
index_upper <- which(sort(upper_bound_random_child$opt_freq$frequency_exceeding)>= 0.05)[1] 

lower_bound = val(index = index_lower, frequency_exceeding = lower_bound_random_child$opt_freq$frequency_exceeding)
upper_bound = val(index = index_upper, frequency_exceeding = upper_bound_random_child$opt_freq$frequency_exceeding)

lower_bound
upper_bound
```

