---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ppvsbp

<!-- badges: start -->
<!-- badges: end -->

The aim of `ppvsbp` package is to make available the code used in the paper
" A comparison between precise and bounded probability to quantify epistemic uncertainty
 in a scientific assessment"
...

## Installation

You can install the development version of `ppvsbp` from [github](https://github.com/Iraices/PrecisePvsBoundedP) with:

```{r}
# remotes::install_github('Iraices/PrecisePvsBoundedP')
```

## Example
This examples does the aluminium exposure assessment by precise probability.

First, let us estimate the probability of exceeding the safety threshold for all chidren

```{r}
library(ppvsbp)
## basic example code
TWI_pp <- unc_analysis_assessment(niter_ale = 1000, niter_epi = 1000, threshold = 1,
                                  percentile_ale = NULL,
                                  suff_stat_concentration = data_assessment$log_concentration_ss_data,
                                  suff_stat_consumption = data_assessment$log_consumption_ss_data,
                                  consumption_change_vals_EKE = data_assessment$change_cons$vals,
                                  consumption_change_probs_EKE = data_assessment$change_cons$probs/100,
                                  consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                                  concentration_mu0 = 3.5, concentration_v0 = 5, concentration_alpha0 = 1,
                                  concentration_beta0 = 1, sufficient_statistics_concentration = TRUE,
                                  consumption_mu0 = -3, consumption_v0 = 5, consumption_alpha0 = 1,
                                  consumption_beta0 = 1, sufficient_statistics_consumption = TRUE,
                                  consumption_event_alpha0 = 1, consumption_event_beta0 = 1)
## Outputs
str(TWI_pp)

## Expected probability of exceeding the safety threshold
TWI_pp$expected_prob_exceed

## Highest posterior density interval of the probability of exceeding the safety threshold
TWI_pp$hdi_prob_exceed

```

Second, let us estimate the probability of exceeding the safety threshold for a high consumer child. Note, the argument
`percentile_ale` has changed `percentile_ale = 95`.

```{r}
  TWI_pp_high_consumer <- unc_analysis_assessment(niter_ale = 1000, niter_epi = 1000, threshold = 1,
                                   percentile_ale = 95,
                                   suff_stat_concentration = data_assessment$log_concentration_ss_data,
                                   suff_stat_consumption = data_assessment$log_consumption_ss_data,
                                   consumption_change_vals_EKE = data_assessment$change_cons$vals,
                                   consumption_change_probs_EKE = data_assessment$change_cons$probs/100,
                                   consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                                   concentration_mu0 = 3.5, concentration_v0 = 5, concentration_alpha0 = 1,
                                   concentration_beta0 = 1, sufficient_statistics_concentration = TRUE,
                                   consumption_mu0 = -3, consumption_v0 = 5, consumption_alpha0 = 1,
                                   consumption_beta0 = 1, sufficient_statistics_consumption = TRUE,
                                   consumption_event_alpha0 = 1, consumption_event_beta0 = 1)

## Outputs
str(TWI_pp_high_consumer)

## Expected probability of exceeding the safety threshold
TWI_pp_high_consumer$expected_prob_exceed

## Highest posterior density interval of the probability of exceeding the safety threshold
TWI_pp_high_consumer$hdi_prob_exceed
```

We can now visualize the results. Epistemic uncertainty has been characterized by a full probability distribution

```{r}
 ## All children
  graph_pp(prob_exceed = TWI_pp$prob_exceed)
    
 ## A high consumer child
  graph_pp(prob_exceed = TWI_pp_high_consumer$prob_exceed)
    
 ## both graphs
   # graph_pp_both(prob_exceed_all = TWI_pp$prob_exceed, 
   #                prob_exceed_high_consumer = TWI_pp_high_consumer$prob_exceed)
```

Now, let us estimate the probability of exceeding the safety threshold by bounded probability. 

First, let us consider all children

```{r}
  ## All children
  lower_bound_prob <- bound_prob_exceed_bp(obj_func_bp = obj_func_bp, maximize = FALSE, 
                                            lower_parameters  = c(1, -5, -20), 
                                            upper_parameters  = c(6, 1, -10),
                                            niter_ale = 1000, niter_epi = 1000, threshold = 1, percentile_ale = NULL,
                                            suff_stat_concentration = data_assessment$log_concentration_ss_data,
                                            suff_stat_consumption = data_assessment$log_consumption_ss_data,
                                            consumption_change_vals_EKE = c(-15, 7.5), 
                                            consumption_change_probs_EKE = c(0.25, 0.75),
                                            consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                                            concentration_mu0 = 2.75,
                                            concentration_v0 = 5, concentration_alpha0 = 1, concentration_beta0 = 1, 
                                            sufficient_statistics_concentration = TRUE,
                                            consumption_mu0 = -2.5,
                                            consumption_v0 = 5, consumption_alpha0 = 1, consumption_beta0 = 1, 
                                            sufficient_statistics_consumption = TRUE,
                                            consumption_event_alpha0 = 1, consumption_event_beta0 = 1)
  

  
  ## All children
  upper_bound_prob <- bound_prob_exceed_bp(obj_func_bp = obj_func_bp, maximize = TRUE, 
                                           lower_parameters  = c(1, -5, -20), 
                                           upper_parameters  = c(6, 1, -10),
                                           niter_ale = 1000, niter_epi = 1000, threshold = 1, 
                                           percentile_ale = NULL,
                                           suff_stat_concentration = data_assessment$log_concentration_ss_data,
                                           suff_stat_consumption = data_assessment$log_consumption_ss_data,
                                           consumption_change_vals_EKE = c(-15, 7.5), 
                                           consumption_change_probs_EKE = c(0.25, 0.75),
                                           consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                                           concentration_mu0 = 2.75,
                                           concentration_v0 = 5, concentration_alpha0 = 1, concentration_beta0 = 1, 
                                           sufficient_statistics_concentration = TRUE,
                                           consumption_mu0 = -2.5,
                                           consumption_v0 = 5, consumption_alpha0 = 1, consumption_beta0 = 1, 
                                           sufficient_statistics_consumption = TRUE,
                                           consumption_event_alpha0 = 1, consumption_event_beta0 = 1)
  
```

Now, let us consider a high consumer child

```{r}
  ## A high consumer child
  lower_bound_high_consumer <- bound_prob_exceed_bp(obj_func_bp = obj_func_bp, maximize = FALSE, 
                                                   lower_parameters  = c(1, -5, -20), 
                                                   upper_parameters  = c(6, 1, -10),
                                                   niter_ale = 1000, niter_epi = 1000, threshold = 1, 
                                                   percentile_ale = 95,
                                                   suff_stat_concentration = data_assessment$log_concentration_ss_data,
                                                   suff_stat_consumption = data_assessment$log_consumption_ss_data,
                                                   consumption_change_vals_EKE = c(-15, 7.5), 
                                                   consumption_change_probs_EKE = c(0.25, 0.75),
                                                   consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                                                   concentration_mu0 = 2.75,
                                                   concentration_v0 = 5, concentration_alpha0 = 1, 
                                                   concentration_beta0 = 1, 
                                                   sufficient_statistics_concentration = TRUE,
                                                   consumption_mu0 = -2.5,
                                                   consumption_v0 = 5, consumption_alpha0 = 1, consumption_beta0 = 1, 
                                                   sufficient_statistics_consumption = TRUE,
                                                   consumption_event_alpha0 = 1, consumption_event_beta0 = 1)
  
  
  
   ## A high consumer child
  upper_bound_high_consumer <- bound_prob_exceed_bp(obj_func_bp = obj_func_bp, maximize = TRUE, 
                                                    lower_parameters  = c(1, -5, -20), 
                                                    upper_parameters  = c(6, 1, -10),
                                                    niter_ale = 1000, niter_epi = 1000, threshold = 1, 
                                                    percentile_ale = 95,
                                                    suff_stat_concentration = data_assessment$log_concentration_ss_data,
                                                    suff_stat_consumption = data_assessment$log_consumption_ss_data,
                                                    consumption_change_vals_EKE = c(-15, 7.5), 
                                                    consumption_change_probs_EKE = c(0.25, 0.75),
                                                    consumers_info_sample_size = data_assessment$consumers_info_sample_size,
                                                    concentration_mu0 = 2.75,
                                                    concentration_v0 = 5, concentration_alpha0 = 1, 
                                                    concentration_beta0 = 1, 
                                                    sufficient_statistics_concentration = TRUE,
                                                    consumption_mu0 = -2.5,
                                                    consumption_v0 = 5, consumption_alpha0 = 1, consumption_beta0 = 1, 
                                                    sufficient_statistics_consumption = TRUE,
                                                    consumption_event_alpha0 = 1, consumption_event_beta0 = 1)
  

```

We can visualize the results individualy or together. In this case, epistemic uncertainty has been characterized by a p-box.

Individual figures

```{r}
 ## All children
 graph_bp(lower_points = lower_bound_prob$opt_prob$prob_exceed, upper_points = upper_bound_prob$opt_prob$prob_exceed)
    
 ## A high consumer child
 graph_bp(lower_points = lower_bound_high_consumer$opt_prob$prob_exceed, 
          upper_points = upper_bound_high_consumer$opt_prob$prob_exceed)
```

Both figures together

```{r}  
 ## Both graphs
  graph_bp_both(lower_points_all = lower_bound_prob$opt_prob$prob_exceed, 
                upper_points_all = upper_bound_prob$opt_prob$prob_exceed,
                lower_points_high_consumer = lower_bound_high_consumer$opt_prob$prob_exceed, 
                upper_points_high_consumer = upper_bound_high_consumer$opt_prob$prob_exceed)
```

End

